\textsf{Are large language models superhuman chemists?}

### 2024-03-21 src/tex/appendix.tex
update manuscript

@@

@@

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/human_score_distribution.pdf}
    \script{plot_human_score_distribution.py}
    \label{fig:human_score_distribution}
    \caption{\textbf{Distribution of human scores.} The histogram shows the fraction of questions that have been answered completely correctly.
    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
\end{figure}

@@

Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), \(p \approx \variable{output/spearman_experience_score_p.txt}\)).

>>>>

\begin{figure}[htb]
    \centering
    \includegraphics{figures/parallel_coordinates_human_aligned.pdf}
    \caption{\textbf{Performance of the models on the different topics of the \enquote{tiny} subset.} The parallel coordinates plot shows the performance of the models on the different topics of the \enquote{tiny} subset. The performance is measured as the fraction of questions that were answered completely correctly by the models. }
    \label{fig:parallel_coordinates_human_aligned}
    \script{make_parallel_coordinates_plot.py}
\end{figure}

@@

\Cref{fig:performance_per_topic_tiny} shows the same analysis on the \enquote{tiny} subset.

@@

One might wonder if questions that are more difficult to parse lead to worse performance of the models.
\Cref{fig:reading_ease_vs_model_performance} shows that there is no clear correlation between the reading ease of the questions and the performance of the models.

@@

\paragraph{Statistics}
\Cref{fig:human_score_distribution} shows the distribution of scores our human scorers achieved.

\begin{figure}[htb]
    \centering
    \includegraphics{figures/human_score_distribution.pdf}
    \script{plot_human_score_distribution.py}
    \label{fig:human_score_distribution}
    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions that have been answered completely correctly.
    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
\end{figure}

We also recorded the time humans took to answer the questions. This time is the time from the question being displayed to the human to the human submitting the answer.
Interestingly, we found no significant correlation between the experience of the human scorers and the performance on the questions (\Cref{fig:human_timing}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), \(p \approx \variable{output/spearman_experience_score_p.txt}\)).

@@

Additionally, we prompted users to provide additional information about their experience in chemistry. 
While we recorded very fine-grained information, e.g., including their specialization, we focussed our analysis on the number of years since the first university-level chemistry course.
\Cref{fig:experience_vs_correctness} shows that the experience of the human scorers was not significantly correlated with the correctness of their answers (\Cref{fig:experience_vs_correctness}, Spearman's \(\rho \approx \variable{output/spearman_experience_score.txt}\), \(p \approx \variable{output/spearman_experience_score_p.txt}\)).

### 2024-03-21 src/tex/ms.tex
human score distribution rule
update manuscript

\paragraph{\enquote{Tiny} subset}
It is important to note that for routine evaluations, a smaller subset of the corpus might be more practical.\autocite{polo2024tinybenchmarks}
For instance,~\textcite{liang2023holistic} report costs of more than 10,000 USD for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark. 
To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus (see \Cref{sec:subset-selection} for details on the curation process).
We also used this subset to seed the web application we used for the human baseline study.

@@

To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar}. 
In this figure, we show the percentage of questions that the models answered completely correctly.
We use horizontal bars to indicate the performance of various models and highlight statistics (e.g., the average) of the human performance with vertical lines.
Remarkably, the figure shows that the leading \gls{llm}, GPT-4, outperforms the best human in our study in this overall metric and vastly, by more than a factor of two, exceeds the average performance of the experts in our study.
Many other models also outperform the average human performance. It is, however, interesting to note that the Galactica model that was trained specifically for scientific applications is outperformed by more general models.
% Bar chart

@@

\paragraph{Question selection} \label{sec:subset-selection}
Since we anticipated that we will not be able to collect enough responses for every question to allow for a meaningful statistical analysis, we decided on showing a relevant subset of all questions to the human scorers.
For selecting the subset, we decided on addressing two questions:
\begin{itemize}
    \item Are the questions for which the models scored poorly just too difficult or unanswerable?
    \item Are there areas in which the performance of humans is very different from the ones of the models?
\end{itemize}
To answer the first question we selected up to 13 questions per source that all \glspl{llm} (model names) from an initial scoring round did not answer correctly.
In addition, we picked 100 diverse one using greedy MaxMin sampling on the embeddings on the questions computed using BART and added five random questions about the number of \gls{nmr} peaks and the point group of molecules.

>>>>

\paragraph{\enquote{Tiny} subset}
It is important to note that for routine evaluations, a smaller subset of the corpus might be more practical.\autocite{polo2024tinybenchmarks}
For instance,~\textcite{liang2023holistic} report costs of more than 10,000 USD for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark. 
To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus in which topics are balanced than in the complete corpus (see \Cref{sec:subset-selection} for details on the curation process).
We also used this subset to seed the web application we used for the human baseline study.

@@

To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar}. 
In this figure, we show the percentage of questions that the models answered completely correctly.
Moreover, we show the worst, best, and average performance of the human experts in our study which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
Remarkably, the figure shows that the leading \gls{llm}, GPT-4, outperforms the best human in our study in this overall metric and vastly, by more than a factor of two, exceeds the average performance of the experts in our study.
Many other models also outperform the average human performance. It is, however, interesting to note that the Galactica model that was trained specifically for scientific applications is outperformed by more general models.
% Bar chart

@@

\paragraph{Question selection} \label{sec:subset-selection}
Since we anticipated that we will not be able to collect enough responses for every question to allow for a meaningful statistical analysis, we decided on showing a relevant subset of all questions to the human scorers.
For selecting the subset, we decided on addressing two questions:
\begin{itemize}
    \item Are the questions for which the models scored poorly just too difficult or unanswerable?
    \item Are there areas in which the performance of humans is very different from the ones of the models?
\end{itemize}
To answer the first question we selected up to 13 questions per source that all \glspl{llm} (model names) from an initial scoring round did not answer correctly.
In addition, we picked 100 diverse one using greedy MaxMin sampling on the embeddings on the questions computed using BART and added five random questions about the number of \gls{nmr} peaks and the point group of molecules.
% https://github.com/lamalab-org/chem-bench-app/issues/92 has the details
